---
title: "ngrams"
author: "Kara C. Hoover"
date: "6/15/2021"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=FALSE, echo = FALSE, warning = FALSE, message = FALSE)
library(stringr)
library(dplyr)
library(data.table)
library(tidytext)
```

## Data Preparation
Sample data, 40%
Course for data is found here:
https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip

Source for bad words list is found here:
bw <- "https://www.freewebheaders.com/download/files/base-list-of-bad-words_csv-file_2021_01_18.zip"


```{r samples, warning=FALSE, message=FALSE}
#read data
blogs <- readLines("final/en_US/en_US.blogs.txt", skipNul=TRUE,
  encoding = "UTF-8")
news <- readLines("final/en_US/en_US.news.txt", skipNul=TRUE,
  encoding = "UTF-8")
twitter<-readLines("final/en_US/en_US.twitter.txt", skipNul=TRUE,
  encoding = "UTF-8")

#dataframes for manipulation
blogs   <- tibble(text = blogs)
news    <- tibble(text = news)
twitter <- tibble(text = twitter)

#sample
set.seed(2021) 
samplesize <- .40 #set sample size

blogs.sample <- blogs %>%
  sample_n(., nrow(blogs)*samplesize)
news.sample <- news %>%
  sample_n(., nrow(news)*samplesize)
twitter.sample <- twitter %>%
  sample_n(., nrow(twitter)*samplesize)

#combine samples
text.sample <- bind_rows(
  mutate(blogs.sample, source="blogs"),
  mutate(news.sample, source="news"), 
  mutate(twitter.sample, source="twitter"))
text.sample$source <- as.factor(text.sample$source)
text.sample <- tibble(text.sample)

#tidy
rm(blogs, news, twitter,
  blogs.sample, news.sample, twitter.sample,
  samplesize)
```

## Clean Text Data
The ngram generation process makes data to lower case so that step is omitted in data cleaning


#remove symbols 
If encoding character remain, try
iconv(text.sample, from = 'UTF-8', to = 'ASCII//TRANSLIT')
save(text.sample, file = "text-ascii.RData", ascii=TRUE)

```{r clean, message=FALSE, results='hold', warning=FALSE}
#remove punct except apost, symbols, http, www
text.clean <- text.sample %>%
  mutate(text=str_replace_all(text, "[^[:alnum:][:space:]']", "")) %>%
  mutate(text=str_replace_all(text, "[[:digit:]]", "")) %>%
  mutate(text=str_replace_all(text, " ?(f|ht)tp(s?)://(.*)[.][a-z]+", "")) %>%
  mutate(text=str_replace_all(text, " www\\.[^ ]*", ""))

text.clean <- rename(text.clean, word = text)

#tidy
rm(text.sample)
```

## Remove profanity
```{r profanity}
#get badwords file
badwords <- read.csv(
    "base-list-of-bad-words_csv-file_2021_01_18.csv", 
    skip=12, skipNul=TRUE, stringsAsFactors = FALSE, header=FALSE,
  col.names =c("word", "v1", "v2", "v3"), na.strings =
    "-------------------------------------------------------")
badwords <- data.frame(badwords)
badwords <- select(badwords, "word")

#remove words
text.clean <- text.clean %>% anti_join(badwords)

#tidy
rm(badwords); gc()
```

## Create ngrams 
```{r ngrams}
unigrams <- text.clean %>%
  unnest_tokens(unigram, word, token = "ngrams", n = 1) %>%
  count(unigram, sort = TRUE)
bigrams <- text.clean %>%
  unnest_tokens(bigram, word, token = "ngrams", n = 2) %>%
  count(bigram, sort = TRUE)
trigrams <- text.clean %>%
  unnest_tokens(trigram, word, token = "ngrams", n = 3) %>%
  count(trigram, sort = TRUE)
quadgrams <- text.clean %>%
  unnest_tokens(quadgram, word, token = "ngrams", n = 4) %>%
  count(quadgram, sort = TRUE)
```

## Clean up
```{r rmClean}
rm(text.clean)
gc()
```

##  checkpoint: save ngrams
Save so that debugging can start with ngrams and save time re-generating.

```{r saveNgrams}
fwrite(unigrams, "unigrams.txt")
fwrite(bigrams, "bigrams.txt")
fwrite(trigrams, "trigrams.txt")
fwrite(quadgrams, "quadgrams.txt")
```

## checkpoint read ngrams
If debugging, start here with saved ngrams to save time regenerating.

```{r readNgrams}
unigrams <- fread("unigrams.txt", data.table=FALSE)
bigrams <- fread("bigrams.txt", data.table=FALSE)
trigrams <- fread("trigrams.txt", data.table=FALSE)
quadgrams <- fread("quadgrams.txt", data.table=FALSE)
```

## Trim ngrams by proporation
ngrams from 28June21 had count over 10 and 90% coverage. Here, the plus 10 method trimmed all values out. reverted to prop90 even though bigger in size.

```{r trimNgrams}
#percent
prop90.unigram <- unigrams %>% count(unigram) %>%  
  mutate(proportion=n / sum(n)) %>% 
  arrange(desc(proportion)) %>%
  mutate(coverage = cumsum(proportion))%>%
  filter(coverage <= 0.9)
prop90.bigram <- bigrams %>% count(bigram) %>%  
  mutate(proportion=n / sum(n)) %>% 
  arrange(desc(proportion)) %>%
  mutate(coverage = cumsum(proportion))%>%
  filter(coverage <= 0.9) 
prop90.trigram <- trigrams %>% count(trigram) %>%  
  mutate(proportion=n / sum(n)) %>% 
  arrange(desc(proportion)) %>%
  mutate(coverage = cumsum(proportion))%>%
  filter(coverage <= 0.9)
prop90.quadgram <- quadgrams %>% count(quadgram) %>% 
  mutate(proportion=n / sum(n)) %>% 
  arrange(desc(proportion)) %>%
  mutate(coverage = cumsum(proportion))%>%
  filter(coverage <= 0.9)
```

## Checkpoint: Write prop filtered data and tidy
If debugging, save data here

```{r filterData}
fwrite(prop90.unigram, "unigrams.trim")
fwrite(prop90.bigram, "bigrams.trim")
fwrite(prop90.trigram, "trigrams.trim")
fwrite(prop90.quadgram, "quadgrams.trim")

rm(unigrams, bigrams, trigrams, quadgrams,
  prop90.unigram, prop90.bigram, prop90.trigram, prop90.quadgram)
gc()
```


## Checkpoint: Read data
If debugging, read data back here.

```{r readFiltered}
#read data
prop90.unigram <- fread("unigrams.trim")
prop90.bigram <- fread("bigrams.trim")
prop90.trigram <- fread("trigrams.trim")
prop90.quadgram <- fread("quadgrams.trim")
```


# Index
```{r index}
setkey(setDT(prop90.bigram), word1, word2)
setkey(setDT(prop90.trigram), word1, word2, word3)
setkey(setDT(prop90.quadgram), word1, word2, word3, word4)
```

## split into separate words 
```{r separate}
#separate ngrams
bigrams.sep <- prop90.bigram %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>% na.omit()
trigrams.sep <- prop90.trigram %>%
  separate(trigram, c("word1", "word2", "word3"), sep = " ") %>% na.omit()
quadgrams.sep <- prop90.quadgram %>%
  separate(quadgram, c("word1", "word2", "word3", "word4"), sep = " ") %>% na.omit()
```


## Checkpoint: Save separated ngram data
For debugging, save point.

```{r saveSeparated}
fwrite(bigrams.sep, "sep90.bigrams.txt")
fwrite(trigrams.sep, "sep90.trigrams.txt")
fwrite(quadgrams.sep, "sep90.quadgrams.txt")

rm(sep90.bigrams, sep90.trigrams, sep90.quadgrams)
gc()
```

## Unite ngrams
```{r uniteNgras}
bigrams <- bigrams %>%
  unite(bigram, word1, word2, sep = " ")
trigrams <- trigrams %>%
  unite(trigram, word1, word2, word3, sep = " ")
quadgrams <- quadgrams %>%
  unite(quadgram, word1, word2, word3, word4, sep = " ")
```

## Save united ngram data
```{r save sep ngrams}
fwrite(unigrams, "unigrams.final.txt")
fwrite(bigrams, "bigrams.final.txt")
fwrite(trigrams, "trigrams.final.txt")
fwrite(quadgrams, "quadgrams.final.txt")
```

## Model
use log and add probabilities for speed
looks like standford model takes bigram counts divided against unigram counts
prediction model then is full sentence divided into bigrams with prob added

Minimize perplexity by allowing for unseen words makes better low (low perplexS)
Perplexity is inverse probability of test set (max prob), normalized by number of words

LaPlace and Kneser-Ney smoothing might not be good for these big datasets. Stupid back-off then with no discounting and only relative frequencies. Good turing might be good b/c it predicts novel words more.

## Test Model
use each model for a task and measure accuracy and speed